---
title: "US_Household_Income"
author: "Naeem Chowdhury, Rosa Gradilla"
date: "December 2, 2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

## Summary

### About the data:
The dataset originally developed for real estate and business investment research. Income is a vital element when determining both quality and socioeconomic features of a given geographic location. The following data was derived from over +36,000 files and covers 348,893 location records.

### Household & Geographic Statistics:
- Mean Household Income (double)
- Median Household Income (double)
- Standard Deviation of Household Income (double)
- Number of Households (double)
- Square area of land at location (double)
- Square area of water at location (double)

### Geographic Location:

- Longitude (double)
- Latitude (double)
- State Name (character)
- State abbreviated (character)
- State_Code (character)
- County Name (character)
- City Name (character)
- Name of city, town, village or CPD (character)
- Primary, Defines if the location is a track and block group.
- Zip Code (character)
- Area Code (character)



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=70))
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```


## libraries

```{r libraries}
#Load in needed libraries 
# Run these if you don't have these libraries yet
# install.packages("tidyverse") 
# install.packages("data.table")
# install.packages("plotly") #For pretty plots
library(tidyverse) # For data processing and visualization
library(data.table) # For creating tables
library(plotly)
library(ggplot2)
library(reshape2)
spot_color = "#9ecae1" 
text_color = "#525252"
```

## File management

```{r file_management}
#Create variables for directories
home.dir <- getwd()
data.dir <- './rawData'
viz.dir <- './dataViz'
output.dir <- './output'
```

## Preprocessing

```{r}
#import
df_income <- read_csv(file.path(home.dir, data.dir, 'kaggle_income.csv'), na = c(''))

df_income2 <- df_income %>% filter(Median != 300000 & Median != 0)
```

## Exploration (Playground)


### Histograms
Before subsetting our data, we have a strange bimodal distribution of values. There is a second peak all concentrated at an income of $300k. 
```{r}
# Distribution
ggplot(data = df_income)+
  geom_histogram(mapping = aes(x = df_income$Median, fill = df_income$Type), bins = 40) +
  labs(x = "Household Income (Median)") +
  scale_fill_discrete(name = "Household Type")
```

Afterwards, we have a much more natural distribution.
```{r}
#number 1
# These histograms satisfy the distribution requirement of the project.
ggplot(data = df_income2)+
  geom_histogram(mapping = aes(x = df_income2$Mean, fill = df_income2$Type), bins = 40) +
  labs(x = "Household Income (Mean)") +
  scale_fill_discrete(name = "Household Type")
```

### What's going on with all the rich people?
There are 4k+ data points which have precisely 300000 dollars as their median income value. This is a data anomaly that messes with our summary statistics, so we decided it would be best to toss them out.
```{r}
# There are 4k+ data points which have precisely 300000 dollars as their median income value. This is a data anomaly that messes with our summary statistics, so we decided it would be best to toss them out.
rich_people <- df_income %>% filter(Median == 300000)

count(rich_people)
```


## Analysis

### Correlation Matrix
How are the quantitative variables correlated with each other? Since the bulk of our quantitative variables are summary statistics gathered manually about each location, they will naturally have a high correlation with each other. 

Naturally, we might ask, do the land area and water area of a location have any relationship with the income of its residents?
```{r}
quant_vars <- df_income2 %>% 
  select(Mean, Median, Stdev, AWater, ALand)

cor_quant <- cor(quant_vars)

melted_quant <- melt(cor_quant)

ggplot(data = melted_quant, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```

### Scatter Plots and Linear Regressions
At first glance, there doesn't seem to be much of a relationship between the physical area of a city and the income of the residents. What might happen if we remove outlier values?
```{r}
# An initial scatter plot to see the shape of the relationship
ggplot(quant_vars, aes(x=Median, y=AWater)) +
  geom_point() + 
  labs(title="Median Income vs. Water Area Scatter Plot", x="Median Income", y = "Square Area of Water at Location")
ggplot(quant_vars, aes(x=Median, y=ALand)) +
  geom_point() + 
  labs(title="Median Income vs. Land Area Scatter Plot", x="Median Income", y = "Square Area of Land at Location")

```
It seems that the correlations we produced in the previous section are a result of outlier values that skew the data. Lets to find out what the outliers are.


```{r}
ggplot(quant_vars, aes(x = "", y = AWater)) +
  geom_boxplot() +
  labs(title="Water Area Box Plot", x="", y = "Square Area of Land at Location") +
  coord_flip() +
  theme(plot.title = element_text(color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1))

AWater_stats <- boxplot.stats(quant_vars$AWater)
AWater_stats[1]

ALand_stats <- boxplot.stats(quant_vars$ALand)
ALand_stats[1]

quant_vars.noOutliers <- quant_vars %>% 
  filter( (AWater < 1309742) & (ALand > 40326) & (ALand < 82994663) )

ggplot(quant_vars.noOutliers, aes(x = "", y = ALand)) +
  geom_boxplot() +
  labs(title="Water Area Box Plot", x="", y = "Square Area of Water at Location") +
  coord_flip() +
  theme(plot.title = element_text(color="#666666", face="bold", size=12, hjust=.5),
              axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
# An initial scatter plot to see the shape of the relationship
ggplot(quant_vars.noOutliers, aes(x=Median, y=AWater)) +
  geom_point() + 
  labs(title="Median Income vs. Water Area Scatter Plot", x="Median Income", y = "Square Area of Water at Location") + 
  geom_smooth(method='lm')
ggplot(quant_vars.noOutliers, aes(x=Median, y=ALand)) +
  geom_point() + 
  labs(title="Median Income vs. Land Area Scatter Plot", x="Median Income", y = "Square Area of Land at Location") +
  geom_smooth(method='lm')

```


### Correlation matrix without outliers

```{r}
cor_quant <- cor(quant_vars.noOutliers)

cor_quant

melted_quant <- melt(cor_quant)

ggplot(data = melted_quant, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

```


## One-Sample Statistical Inference for Income

Confidence interval for income's population mean income value.
```{r}
#-Construct a confidence interval for income's population mean value. (Make sure to check assumptions and interpret the CI)
#-Conduct a hypothesis test against a certain value. Interpret its results

#number 2
clevel <- 0.95
n <- length(df_income2$Median)
  
xbar <- mean(df_income2$Median)
s <- sd(df_income2$Median)
se <- s/(sqrt(n))
x <- (1-clevel)/2
p <- clevel + x
cv <- qt(p, n-1)
CI <- c(xbar - (cv*se), xbar + (cv*se))

print(CI)


#Hypothesis Test
#Ho: mu = 50,000      #Ha: mu != 50,000
t.test(df_income2$Median, mu=50000, alternative = "two.sided")
```
We tested against the null hypothesis mu = 50,000 on a two sided t test. Our confidence interval for population's median income is [53145.03, 53812.59]. We are 95% confident the true population's median income falls bewteen our confidence interval 95% of the time. Given that our data came from a random sample and assuming the population's distribution is approximately normal.

We came up with a null hypothesis claiming the true population's median income is equal to 50,000, our t-test against this null hypothesis gave us a p-value < 2.2e-16 thus we reject the null hypothesis. This agrees with our confidence interval given that 50,000 is not part of it.

```{r}
#Cohen's d
d <- abs(xbar - 50000)/s
d

```
Our effect size is 0.1222492. We witness a relatively small practical effect size.

## Categorical Explanatory Variable

Here we are performing a proportion test for the proportion of median income being lower than 50,000
```{r}
#number 3
cuts <- quantile(df_income2$Median, c(0, 1/2, 1))

df_categ <- df_income2 %>% mutate(category=cut(Median, breaks=cuts, labels=c("below","above")))

shaved_income <- df_categ %>% select(category)
table(shaved_income)

prop.table <- table(shaved_income)*(1/27924)

# H_0: The proportion of areas with median income below $50k is 0.5.
# H_a: The proportion of areas with median income below $50k is greater than or less than 0.5.

prop.test(13954, 27924, 0.5)
```
Our confidence interval is [0.4938316, 0.5055955]. We tested a null p equal to 0.5. We failed to reject the null hypothesis. A p-value of 0.9285 indicates that the probability of observing our test statistic p = 0.4997135 is 93% given the null-hypothesis is true.

```{r}
#Cohen's h
x <- asin(sqrt(0.4997135)) - asin(sqrt(0.5))
effect.size <- 2*abs(x)
```
Our effect size is 0.000573, we witness a very small practical effect size.

### Two-sample statistical inference
Two sample: compare means of two population groups: cities and towns

Is there a significant difference between the household incomes in cities and towns?

Ho: mt = mc  i.e. there is no difference
Ha: mt != mc

```{r}
#number 4
#Check assumptions, interpret the results.

city <- df_income2 %>%
    filter(Type == 'City')

town <- df_income2 %>%
  filter(Type == 'Town')

t.test(city$Median, town$Median)
```
95% confidence interval is [747.1329, 5865.2244]. P-value is equal to 0.0114, at a 0.05 significance level we fail to reject the null hypothesis: the true difference in means is not equal to 0. This agrees with our conficence inteval since 0 is not within it.

Checking assumptions:
```{r}
hist(city$Mean)
hist(town$Mean)
```
Since both of our groups seem to be slightly skwed, out t-test might not be reliable since it assumes the data is normally distributed.

Effect size:
```{r}
s1 <- sd(town$Median)
s2 <- sd(city$Median)
effect.size <- abs(47112.28 - 43806.10)/sqrt((s1^2 + s2^2)/2)
```
Our effect size is 0.148239. thus we witness a relatively small effect size. 


For the confidence interval and hypothesis test conducted, we ran simulations: 
we took a random subsample of fixed size (n=100) from our full data set, we then proceeded to calculate the confidence intervals and hypothesis tests against the true sample mean, recording the coverage.

```{r}
#number 5
x <- df_income2$Median
mu <- mean(x)

n.trials <- 10000
n <- 10
inside <-  numeric(n.trials)
for (j in 1:n.trials){
  x.samp <- sample(x, n)
  x.bar <- mean(x.samp)
  x.sd <- sd(x.samp)
  CI[j] <- (mu > x.bar - qt(0.95,n-1)*x.sd/sqrt(n) & mu < x.bar + qt(0.95,n-1)*x.sd/sqrt(n))
  inside[j] <- CI[j]
}

coverage10 <- mean(inside)
cat("For a subsample of 10 points:", coverage10)
print(" ")

n.trials <- 10000
n <- 50
inside <-  numeric(n.trials)
for (j in 1:n.trials){
  x.samp <- sample(x, n)
  x.bar <- mean(x.samp)
  x.sd <- sd(x.samp)
  CI[j] <- (mu > x.bar - qt(0.95,n-1)*x.sd/sqrt(n) & mu < x.bar + qt(0.95,n-1)*x.sd/sqrt(n))
  inside[j] <- CI[j]
}

coverage50 <- mean(inside)
cat("For a subsample of 50 points:", coverage50)
print(" ")

n.trials <- 10000
n <- 100
inside <-  numeric(n.trials)
for (j in 1:n.trials){
  x.samp <- sample(x, n)
  x.bar <- mean(x.samp)
  x.sd <- sd(x.samp)
  CI[j] <- (mu > x.bar - qt(0.95,n-1)*x.sd/sqrt(n) & mu < x.bar + qt(0.95,n-1)*x.sd/sqrt(n))
  inside[j] <- CI[j]
}

coverage100 <- mean(inside)
cat("For a subsample of 100 points:", coverage100)
print(" ")

n.trials <- 10000
n <- 1000
inside <-  numeric(n.trials)
for (j in 1:n.trials){
  x.samp <- sample(x, n)
  x.bar <- mean(x.samp)
  x.sd <- sd(x.samp)
  CI[j] <- (mu > x.bar - qt(0.95,n-1)*x.sd/sqrt(n) & mu < x.bar + qt(0.95,n-1)*x.sd/sqrt(n))
  inside[j] <- CI[j]
}

coverage1000 <- mean(inside)
cat("For a subsample of 1000 points:", coverage1000)
```
Our coverage for the true sample mean over 10,000 trials was 89.60%, a little lower than expected (95%), this could be attributed to outliers in the data. 


## Bootstrap Simulations 

Based on our previous analysis of the quantitative variables, it seems that income is most highly correlated with the water area at the location. We proceed by computing 10,000 bootstrapped simulations of the data. 

```{r}
plot(quant_vars.noOutliers$Mean, quant_vars.noOutliers$AWater, 
     col = "red", pch = 19,
     xlab = "Median Income", ylab = "Area of Water in Location",
     main = "Bootstrapped Linear Model for Income vs. Water")



waterfit <- lm(Median ~ AWater, data = quant_vars.noOutliers)

index <- 1:nrow(quant_vars.noOutliers)
set.seed(7)
n.trials <- 10000
for(i in 1:n.trials){
  index.boot <- sample(index, replace = TRUE) #getting a bootstrap sample of indices
  quant_vars.boot <- quant_vars.noOutliers[index.boot, ] #bootstrapped dataset
  fit.boot <- lm(Median ~ AWater, data = quant_vars.boot)
  abline(fit.boot, lwd = 0.01, col= rgb(0,0.1,0.25, alpha = 0.05)) # add a gray line at that point
  
}
abline(waterfit, lwd=2, col = "blue")
```


